---
title: "Big Five-Cog ctsem"
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

# Workspace

## Packages

```{r}
library(psych)
library(plyr)
library(tidyverse)

```

```{r}
sc_df <- read_csv("Data/ema_wide_clean_data_W1_2025-04-30.csv") 
colnames(sc_df)
```

# EMA Data 

1. Delete EACNO columns; they are the averages of observed values but we need to impute facets then calculate EACNO averages
2. Create StartDate and HourBlock
```{r}
sc_df <- sc_df %>%
  group_by(SID) %>%
  mutate(StartDate = min(Date)) %>%
  ungroup() %>%
  select(SID:interrupt, StartDate, everything())
```


## Big Five States 

```{r}
pp_cb <- readxl::read_excel("Codebooks/01-codebook.xlsx", sheet = "codebook")
b5_items <- pp_cb %>% filter(Inventory == "BFI-2") %>% pull(old)
# sc_df <- sc_df %>%
#   #rename(N05 = N5) %>%
#   select(SID, Date = Date_main, session:HourBlock, interrupt, all_of(b5_items)) %>%
#   pivot_longer(
#     cols = all_of(b5_items)
#     , values_to = "value"
#     , names_to = "old"
#     , values_drop_na = T
#   ) %>%
#   left_join(pp_cb %>% select(old, shortFacet, Reverse)) %>%
#   mutate(value = ifelse(Reverse == "yes", 6 - value, value)) %>%
#   separate(old, c("trait", "item"), sep = -2) %>%
#   group_by(SID, Date, StartDate, Day, Hour, HourBlock, session, interrupt, trait, shortFacet) %>%
#   summarize(value = mean(value, na.rm = T)) %>% 
#   ungroup()

#pomp <- function(x, mini = 1, maxi = 5) (x - mini)/(maxi - mini)*100

sc_b5 <- sc_df %>% 
  group_by(SID, Date, StartDate, Day, Hour, HourBlock, session,  interrupt, trait) %>% 
  summarize(value = mean(value, na.rm = T)) %>% 
  ungroup() %>%
  pivot_wider(
    names_from = "trait"
    , values_from = "value"
    ) %>%
  full_join(
    sc_b5 %>%
      pivot_wider(
        names_from = c("trait", "shortFacet")
        , values_from = "value"
        )
  ) #%>% 
  # mutate_at(vars(A:N_emoVol), pomp)
```

## Cognitive States 

```{r}
#pomp_obs <- function(x) (x - min(x, na.rm = T))/(max(x, na.rm = T)-min(x, na.rm = T))*100
# sc_cog <- sc_df %>% 
#   select(
#     SID, Date = Date_main, session:HourBlock, interrupt, 
#          , nback_score, vpa_score, dsm_score
#          , nback_medianRTc, vpa_medianRTc, dsm_medianRTc
#          ) %>%
  # mutate_at(vars(nback_score, vpa_score, dsm_score), pomp_obs) %>%
  # rowwise() %>%
  # mutate(cog_comp = mean(c_across(nback_score:dsm_score), na.rm = T)) %>%
  ungroup() 

sc_b5cog <- sc_b5 %>%
  inner_join(
    sc_cog 
    ) %>%
  select(-StartDate, -HourBlock) %>%
  arrange(SID, Date) %>%
  group_by(SID) %>%
  mutate(
    beep = 1:n()
    , beep_wpc = beep - median(beep)
    , dec_time = hour(Date) + minute(Date)/60 + second(Date)/60/60
    , dec_time_c = dec_time - 12
  )
save(sc_b5cog, file = sprintf("data-pulls/b5-cog-ctsem/b5-cog-ema-%s.RData", Sys.Date()))
write_csv(sc_b5cog, file = sprintf("data-pulls/b5-cog-ctsem/b5-cog-ema-%s.csv", Sys.Date()))
```


```{r}
set.seed(420)
sc_cog %>%
  group_by(SID) %>%
  filter(n() > 40) %>%
  ungroup() %>%
  filter(SID %in% sample(unique(.$SID), 20)) %>%
  ggplot(aes(x = session, y = dsm_score)) + 
    geom_line() + 
    geom_point() + 
    geom_smooth() + 
    facet_wrap(~SID, nrow = 4) + 
    theme_bw()
  
```

# Baseline Data 
## Big Five Traits  

```{r}
bl_df <- read_csv("parsed-data/baseline/baseline_wide_clean_data_W1_2025-04-30.csv") %>%
  select(
    SID, StartDate = date
    , starts_with("demo_eth"),
    , demo_gender, demo_gender_TEXT = demo_gender_3_TEXT, demo_hispanic
    , demo_orientation, demo_orientation_TEXT = demo_orientation_4_TEXT
    , demo_YOB, demo_age
    , demo_education
    , demo_income
    , demo_relationship 
    # , demo_livingSituation
    # , demo_employment 
    # , demo_mom_edu, demo_dad_edu
    , starts_with("BFI")
    # , health_physAct = `health-physAct` 
    # , health_smoke = `health-smoke`
    # , health_height = `health-height`, health_weight = `health-weight`
    # , health_cholesterol = `health-CC_1`
    # , health_hypertension = `health-CC_2`
    # , health_diabetes = `health-CC_3`
    # , health_depression = `health-CC_4`
    # , health_tbi = `health-CC_5`
    # , health_heartDis = `health-CC_6`
    # , health_renal = `health-CC_7`
    # , starts_with("sud")
    ) %>%
  filter(!is.na(SID))
write_csv(bl_df, file = sprintf("data-pulls/b5-cog-ctsem/b5-cog-baseline-%s.csv", Sys.Date()))
save(bl_df, file = sprintf("data-pulls/b5-cog-ctsem/b5-cog-baseline-%s.RData", Sys.Date()))
```

## Ctsem shite

```{r, include = FALSE}
library(rstan)
library(ctsem)
library(tidybayes)
library(furrr)
library(plyr)
library(tidyverse)

# set.seed(420)
# 
# load("Data/sherman_wide_person_centered.RData") # for multilevel; script 0a
# colSums(is.na(sherman_wide_person_centered)) #not too bad
# desc <- sherman_wide_person_centered %>% psych::describe()
```

## Testing

simple ass test

```{r}
model<-ctModel(type='stanct'
            , n.latent = 2, latentNames   = c("E", "vpa_score")
            , n.manifest = 2, manifestNames = c("E", "vpa_score")
            , TIpredNames = c("E_baseline", "age")
            , TDpredNames = c("interrupt", "Hour", "session")
            , LAMBDA=diag(2)
            , MANIFESTMEANS = 0 #fix this for estimation 
            , CINT = matrix(c("mu_trait", "mu_cog"), nrow = 2) #estimate this for equil calculation
            , DRIFT = matrix(c("beta_trait", "beta_trait_on_cog", "beta_cog_on_trait", "beta_cog"), nrow = 2, byrow = T)
            #diagonal for estimation
            , MANIFESTVAR   = matrix(c("mv_trait", 0, 0, "mv_cog"), nrow = 2, byrow = TRUE)
            , DIFFUSION = matrix(c("diff_trait", 0, 0, "diff_cog"), nrow = 2, byrow = TRUE), 
            , T0VAR = matrix(c("var_trait", 0, 0, "var_cog"), nrow = 2, byrow = TRUE) 
            , id            = "id"
            , time          = "time" )

model$pars$indvarying[7:14] <- TRUE #drift and diffusion
model$pars$transform[11] <- "log1p_exp(2 * param) + 1e-10"
model$pars$transform[14] <- "log1p_exp(2 * param) + 1e-10"

model$pars



model_fit <- ctStanFit(datalong = sherman_wide_person_centered, 
                      ctstanmodel = model,
                      optimize = TRUE,
                      priors = TRUE,
                      cores=10
                      # ,
                      # optimcontrol=list(is=TRUE, 
                      #                   finishsamples = 1000, 
                      #                   isloopsize = 2000)
                      )

# Summary of the model
modelsum <- summary(model_fit, digits = 4, parmatrices=TRUE)
modelpar <- modelsum$parmatrices
print(modelsum$parmatrices)

```
All 30 trait-cog pairs

```{r}
# Define traits and cogs
traits <- colnames(sc_df)[4:8] #change col numbers
cogs <- colnames(sc_df)[9:16] #change col numbers

# Create all trait-cog pairs
trait_cog_pairs <- expand_grid(trait = traits, cog = cogs)

fit_bivariate_model <- function(trait, cog) {
  message("Fitting model for:", trait, "and", cog, "\n")

  tryCatch({
    var_names <- c(trait, cog)

    model <- ctModel(
      type = 'stanct',
      n.latent = 2,
      latentNames = var_names,
      n.manifest = 2,
      manifestNames = var_names,
      LAMBDA = diag(2),
      
      TDpred
      
      MANIFESTMEANS = 0,

      CINT = matrix(c(
        paste0("mu_", trait),
        paste0("mu_", cog)
      ), nrow = 2),

      DRIFT = matrix(c(
        paste0("beta_", trait),
        paste0("beta_", trait, "_on_", cog),
        paste0("beta_", cog, "_on_", trait),
        paste0("beta_", cog)
      ), nrow = 2, byrow = TRUE),

      MANIFESTVAR = matrix(c(
        paste0("mv_", trait), 0,
        0, paste0("mv_", cog)), 
        nrow = 2, byrow = TRUE),

      DIFFUSION = matrix(c(
        paste0("diff_", trait), 0,
        0, #lower diagonal f's stuff up
        paste0("diff_", cog)
      ), nrow = 2, byrow = TRUE),

      T0VAR = matrix(c(
        paste0("t0var_", trait), 0,
        0, paste0("t0var_", cog)), 
        nrow = 2, byrow = TRUE),

      id = "id",
      time = "time"
    )

    model$pars$indvarying[7:14] <- TRUE #drift and diffusion
    
    #make diffusion variances reasonable
    model$pars$transform[11] <- "log1p_exp(2 * param) + 1e-10"
    model$pars$transform[14] <- "log1p_exp(2 * param) + 1e-10"

    data_sub <- sherman_wide_person_centered %>%
      select(id, time, all_of(c(trait, cog)))

    fit <- ctStanFit(datalong = data_sub, 
                     ctstanmodel = model, 
                     optimize = TRUE, 
                     priors = TRUE,
                     cores = 2, 
                     verbose = 1
                     )

    fit_summary <- summary(fit, digits = 4, parmatrices = TRUE)

    tibble(
      pair = paste0(trait, "_", cog),
      model = list(model),
      fit = list(fit),
      summary = list(fit_summary)
    )
  }, error = function(e) {
    cat("Error in model for: ", trait, "_", cog, "\n", e$message)  # Log the error
    tibble(
      pair = paste0(trait, "_", cog),
      model = list(NA),
      fit = list(NA),
      summary = list(e$message)
    )
  })
}

# Parallel plan
plan(multisession, workers = 4)

# Fit all models
ml_bivariate_results <- future_pmap_dfr(trait_cog_pairs, fit_bivariate_model, .progress = TRUE)
#check
# ml_bivariate_results$summary[3][[1]]$parmatrices

# Save results
save(ml_bivariate_results, file = "Data/ml_bivariate_results.RData")

# Return to sequential plan
plan(sequential)

q()

```

Extract parameters and find out which didn't converge and reestimate

Yo we good - 5/14/25

```{r}
# #load("Data/ml_bivariate_results.RData")
# 
# ml_bivariate_summary_t0_fixed <- ml_bivariate_results %>%
#   select(pair, summary)
# 
# # Extract the parmatrices from each summary
# ml_bivariate_parmatrices_t0_fixed <- ml_bivariate_summary_t0_fixed %>%
#   mutate(parmatrices = map(summary, ~ .x$parmatrices)) %>%
#   select(pair, parmatrices)
# 
# #Are any estimates hella big; this is  indicative of poor model convergence and I can separately estimate them
# 
# par_check_t0_fixed <- ml_bivariate_parmatrices_t0_fixed %>%
#   mutate(any_mean_gt_20 = map_lgl(parmatrices, ~ any(.x$Mean > 20, na.rm = TRUE)))
# 
# par_large_means_t0_fixed <- par_check_t0_fixed %>%
#   filter(any_mean_gt_20) %>%
#   mutate(high_mean_rows = map(parmatrices, ~ filter(.x, Mean > 20)))
# 
# par_large_unnest_t0_fixed <- par_large_means_t0_fixed %>% unnest(high_mean_rows) %>%
#   mutate(Mean = round(Mean, 3))
# 
# #im so fuckin happy
```


